Converging High-Throughput and High-Performance Computing Paradigms: \\ A Case Study of Using Abstractions to Bridge the Divide

### 2017-03-13 PandaEvolution.tex
Question IV; transition V-VI

\mtnote{Please feel free to add details about backfill as needed.}
\mtnote{Please note: I know the following departs for the current name given to
the PanDA subsystem installed on Titan. I think there is a good reason to call
it a broker instead of a pilot, and I think I explained it in the previous
paragraphs. Please  take this just as a suggestion, something I would like to
discuss in our meeting.} \sergeynote{It's not a pilot in conventional sense. I
call it an agent. It just happened that we were able to use full PanDA pilot's
code base to serve our purposes on Titan. that's why, by inertia, we still call
it a pilot.}\mtnote{Thank you. We would have a problem calling it `agent' as we
use that term to name the pilot of NGE. Would PanDA Broker work?}

@@

Once deployed, every PanDA Broker supports the execution of MC
% event generation and
detector simulation on Titan in 9 steps: (i) PanDA Broker queries Titan's
Moab scheduler about currently unused worker nodes;
% and checks whether the time and size of the available backfill resources are
% suitable to execute jobs;
(ii) the PanDA Broker queries the Job Dispatcher module of the PanDA server for
jobs that have been bound to Titan by JEDI; (iii)  the PanDA Broker requests to
the Data Service module of the PanDA Server what input data these jobs require
and transfer them to the DTN via the ATLAS DDM service; (iv) the PanDA Broker
creates one or more PBS MPI script, wrapping enough jobs' payload to fit
backfill availability; (v) the PanDA Broker submits these MPI jobs to the
Titan's PBS batch system via RADICAL-SAGA; (vi) the MPI jobs waits a minimal
amount of time into the queue until they are scheduled for execution; (vii) upon
execution on the worker nodes, each MPI script creates a ramdisk and sets up the
specific AthenaMP configuration required by the wrapped payload; (viii) the
payload is executed into dedicated processes, 16 for each work node, 1 for each
available core; (ix) upon completion of an MPI job, the PanDA Broker locates the
produced output files on the shared filesystem and transfer them to BNL ATLAS
Tier 1 computing centre.

>>>>

\mtnote{Please feel free to add details about backfill as needed.}
\mtnote{Please note: I know the following departs for the current name given to
the PanDA subsystem installed on Titan. I think there is a good reason to call
it a broker instead of a pilot, and I think I explained it in the previous
paragraphs. Please  take this just as a suggestion, something I would like to
discuss in our meeting.} \sergeynote{It's not a pilot in conventional sense. I
call it an agent. It just happened that we were able to use full PanDA pilot's
code base to serve our purposes on Titan. that's why, by inertia, we still call
it a pilot.}\mtnote{Thank you. We would have a problem calling it `agent' as we
use that term to name the pilot of NGE. Would PanDA Broker work?}\mtnote{Sergey
agreed that we should not use `pilot'. He is quite convinced by `broker' but we
are both open to better suggestions.}

@@

\mtnote{Does one PanDA Broker support the submission and monitoring of one job
at a time? Or does it support the concurrent submission and monitoring of
multiple jobs (assuming enough backfill availability)?}

Once deployed, every PanDA Broker supports the execution of MC detector
simulation on Titan in 9 steps: (i) PanDA Broker queries Titan's Moab scheduler
about currently unused worker nodes; (ii) the PanDA Broker queries the Job
Dispatcher module of the PanDA server for jobs that have been bound to Titan by
JEDI; (iii)  the PanDA Broker requests to the Data Service module of the PanDA
Server what input data these jobs require and transfer them to the DTN via the
ATLAS DDM service; (iv) the PanDA Broker creates one or more PBS MPI script,
wrapping enough jobs' payload to fit backfill availability; (v) the PanDA Broker
submits these MPI jobs to the Titan's PBS batch system via RADICAL-SAGA; (vi)
the MPI jobs waits a minimal amount of time into the queue until they are
scheduled for execution; (vii) upon execution on the worker nodes, each MPI
script creates a ramdisk and sets up the specific AthenaMP configuration
required by the wrapped payload; (viii) the payload is executed into dedicated
processes, 16 for each work node, 1 for each available core; (ix) upon
completion of an MPI job, the PanDA Broker locates the produced output files on
the shared filesystem and transfer them to BNL ATLAS Tier 1 computing centre.

### 2017-03-13 nge.tex
Question IV; transition V-VI

Pilots can offer a payload-independent scheduling interface while hiding the
mechanics of coordination and communication among multiple work nodes. This
could eliminate the need for packaging payload into PBS jobs within the broker,
greatly simplifying the submission process. This simplification would also
enable the submission of different types of payload, without having to develop a
specific PBS script for every payload. Finally, the submission process would be
MPI-independent, as MPI is used for coordination among multiple work nodes, not
by the payload.

\mtnote{NOTE: I do not mention backfill availability Vs normal availability
(i.e., submitting to the queue a `normal' job) because, in theory, the design of
PanDA Broker enables this capability. I would speak about about backfill/not
backfill in the discussion of NGE, when speaking about its generality. Part of
this generality is being able to submit to whatever batch system is supported by
SAGA, including Titan's PBS and all its queues.}

\mtnote{NOTE: NGE is not able to use backfill on Titan as we do not have specific functionalities to interrogate the Moab scheduler about backfill availability. This is why NGE should be presented as the pilot for PanDA Broker and not as an alternative to PanDA Broker. NGE alone would not be able to speak directly to PanDA Server.}

>>>>

Multiple generations would also help with the transition from an execution model
based on a fixed amount of event per simulation (100 at the moment) to one in
which the number of events is decided on the base of the available walltime.
Pilots would enable to package simulations with a homogeneous number of events,
especially when core and walltime availabilities would diverge as when a small
amount of cores is available for a long amount of time or \textit{vice versa}.
In this situations, a PBS job would would require all simulation to have either
a large or a small number of events; a pilot would instead support multiple
generations of simulations with a different number of events and therefore
duration.

Pilots can offer a payload-independent scheduling interface while hiding the
mechanics of coordination and communication among multiple work nodes. This
could eliminate the need for packaging payload into PBS jobs within the broker,
greatly simplifying the submission process. This simplification would also
enable the submission of different types of payload, without having to develop a
specific PBS script for each payload. The submission process would also be
MPI-independent, as MPI is used for coordination among multiple work nodes, not
by the payload.

\mtnote{I tried to argue that pilots facilitates the submission to a normal
queue but I was not able to find a convincing argumentation. When we consider
only Titan, a PBS job is the only way to submit a job to its queue(s). With NGE,
we translate a pilot description to a PBS job and then submit it to a queue via
a PBS command. PanDA Brokers do something very similar: they package detector
simulations into a PBS job and they submit it via a PBS command. The fact that
PanDA Brokers define the number of cores and walltime of a PBS job on the base
of backfill information, does not make the packaging or the submission processes
any different: PanDA Brokers are already submitting to the nomal queue. If I got
it right, PanDA Brokers could already submit a job with, say, 12000 cores and 2
hours walltime. They don't do it because they have no allocation and they would
wait a long time in the queue (maybe, let's see what our experiments say about
that). I do not see how NGE could make all this any different.}

\mtnote{NOTE: I would speak about about backfill/not backfill in the discussion
of NGE, when speaking about its generality towards resources, i.e., unified
submission and scheduling process across different resources and multiple
resources. Part of this generality is being able to submit to whatever batch
system is supported by SAGA (including Titan's PBS and all its queues) and, in
case, to multiple resources at the same time.}

\mtnote{NOTE: NGE is not able to use backfill on Titan as we do not have
specific functionalities to interrogate the Moab scheduler about backfill
availability. This is why NGE should be presented as the pilot for PanDA Broker
and not as an alternative to PanDA Broker. Further, NGE alone would not be able
to speak directly to PanDA Server.}
